---
title: "Practical Machine Learning Course Project"
author: "Hitoshi Kawashima"
date: "February 12, 2017"
output: html_document
---

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Objectives

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.

## Getting and Cleaning Data

```{r}
# Load data files
training <- read.csv("~/Desktop/Coursera/pml-training.csv")
testing <- read.csv("~/Desktop/Coursera/pml-testing.csv")

str(training)
dim(training)
dim(testing)

# Load relevant packages
library(ggplot2)
library(caret)
library(rattle)
library(randomForest)

# Create a training (training1) and validation (training2) subsets within the training data
set.seed(10)
inTrain <- createDataPartition(y = training$classe, p = 0.7, list = FALSE)
training1 <- training[inTrain, ]
training2 <- training[-inTrain, ]
```

### Remove variables that could throw off the model from both the training and validation data sets: a) near zero variance b) mostly NA c) identifiers

```{r}
# a) near zero variance
nzv <- nearZeroVar(training1)
training1 <- training1[, -nzv]
training2 <- training2[, -nzv]

# b) mostly NA
mostlyNA <- sapply(training1, function(x) mean(is.na(x))) > 0.95
training1 <- training1[, mostlyNA == FALSE]
training2 <- training2[, mostlyNA == FALSE]

# c) identifiers, first 5 variables in the data set
training1 <- training1[, -(1:5)]
training2 <- training2[, -(1:5)]
```

## rpart model

```{r}
set.seed(10)

# Fit model on training1
mod_rpart <- train(classe ~ ., method = "rpart", data = training1)

# Print final model
print(mod_rpart$finalModel)

# Create a nicer graphic
fancyRpartPlot(mod_rpart$finalModel)

# Predict class in validation subset using the rpart model
pred_rpart <- predict(mod_rpart, training2)

# Call the confusion matrix to see the accuracy / estimated out-of-sample error
confusionMatrix(training2$classe, pred_rpart)
```
### The accuracy is not that high so try another model

## Random forest model

```{r}
set.seed(10)

# Fit model on training1
mod_rf <- randomForest(classe ~ ., data = training1)

# Print model
print(mod_rf)

# Predict class in validation subset using the random forest model
pred_rf <- predict(mod_rf, training2)

# Call the confusion matrix to see the accuracy / estimated out-of-sample error
confusionMatrix(training2$classe, pred_rf)
```
### The accuracy is 99.75% so the estimated out-of-sample error is 0.25%

## Creating answer files

```{r}
# Predict class in testing set using the random forest model
answers <- predict(mod_rf, newdata = testing)

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_", i ,".txt")
    write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, col.names = FALSE)
  }
}

pml_write_files(answers)
print(answers)
```